---
title: "NYPD_Shooting_Incident"
author: 
date: "2023-11-08"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Source
Here is the source of the data (NYPD Shotting Incident).

https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD

## Importing Data
Here is the raw data before any changes or clean-up.
Also, all the libraries that are needed for this file.
```{r data importing, echo=TRUE, warning=FALSE, message=FALSE}
library(tidyverse)
library(lessR)
url_in <- "https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD"
NYPD_Whole_Data <- read_csv(url_in)
NYPD_Whole_Data
```

## Data Cleaning
There are total 21 columns but I will select some columns that I am interested and
organize them by factoring.
```{r data cleaning, echo=TRUE}
NYPD_My_Interest <- NYPD_Whole_Data %>%
    select(INCIDENT_KEY, OCCUR_DATE, OCCUR_TIME, BORO, PRECINCT, STATISTICAL_MURDER_FLAG,
           PERP_AGE_GROUP, PERP_SEX, PERP_RACE, VIC_AGE_GROUP, VIC_SEX, VIC_RACE) %>%
    mutate(OCCUR_DATE = mdy(OCCUR_DATE),INCIDENT_KEY = as.factor(INCIDENT_KEY),
            BORO = as.factor(BORO), PRECINCT = as.factor(PRECINCT), 
            STATISTICAL_MURDER_FLAG = as.factor(STATISTICAL_MURDER_FLAG),
            PERP_AGE_GROUP = as.factor(PERP_AGE_GROUP), PERP_SEX = as.factor(PERP_SEX),
            PERP_RACE = as.factor(PERP_RACE), VIC_AGE_GROUP = as.factor(VIC_AGE_GROUP),
            VIC_SEX = as.factor(VIC_SEX), VIC_RACE = as.factor(VIC_RACE))
summary (NYPD_My_Interest)
```

## Visualization
Let's see which borough has the highest and lowest shooting incident.
```{r visualization, echo=TRUE}
#Grouping the data by each borough before plotting a graph.
Each_Boro <- NYPD_My_Interest %>% group_by(BORO) %>%
  summarize(incidents = n())

ggplot(Each_Boro, aes(x=BORO, y=incidents, fill=BORO)) +
  geom_bar(stat="identity") +
  xlab("Borough") + ylab("Incidents") +
  ggtitle("Incidents in Different Boroughs") +
  theme_minimal()
```
It looks like Brooklyn has the highest and Staten Island has the least incidents.



Let's visualize the age group of victims.
The reason why I did not choose to visualize any data related to perpetrator is because
there are too many Null and Unknown.
```{r table1, echo=TRUE}
table(NYPD_My_Interest$PERP_AGE_GROUP)
```
```{r table2, echo=TRUE}
table(NYPD_My_Interest$PERP_RACE)
```
```{r table3, echo=TRUE}
table(NYPD_My_Interest$PERP_SEX)
```



Here is the table of victim's age group.
```{r tables, echo=TRUE}
table(NYPD_My_Interest$VIC_AGE_GROUP)
```

The 2nd column says "1022" which seems like a typo and it's just 1 person so I will 
just include that to "UNKNOWN".
```{r changes, echo=TRUE, warning=FALSE, message=FALSE}
NYPD_My_Interest$VIC_AGE_GROUP <- replace(NYPD_My_Interest$VIC_AGE_GROUP, 
                                          NYPD_My_Interest$VIC_AGE_GROUP == 1022, 'UNKNOWN')
VIC_AGE <- NYPD_My_Interest %>%
  select(VIC_AGE_GROUP) %>%
  filter(VIC_AGE_GROUP != 'UNKNOWN')

table(VIC_AGE)

#1022 and UNKNOWN columns still exist so I would like to get rid of them.
VIC_AGE$VIC_AGE_GROUP <- droplevels(VIC_AGE$VIC_AGE_GROUP)


PieChart(VIC_AGE_GROUP, hole = 0, values = "%", data = VIC_AGE, 
         main = "Age Groups of Victims", labels_cex = 1.0, quiet = TRUE)
```
The age group 25-44 has the most victim but 18-24 also has a lot of victims 
in that age group. 


## Data Model
I would like to create a logistic regression model that shows the probability that
murder case relates to some variables like perpetrator's age group, sex and race.
```{r data model, echo=TRUE}
glm.fit <- glm(STATISTICAL_MURDER_FLAG ~ PERP_RACE + PERP_SEX + PERP_AGE_GROUP, 
               data = NYPD_My_Interest, family = binomial)
summary(glm.fit)
```
By looking at Pr(>|z|), which is also called p-value, of each predictor variable,
PER_RACEBLACK(0.0714) and PER_RACKBLACK HISPANIC(0.0336) are statistically significant
for the prediction of if an shooting incident is murder or not. The significance level can 
be determined by looking at the p-value. If the p-value is lower than 0.05 it is highly
significant and if it is lower than 0.10, I would still consider that as significant.

The p-values of PER_SEX are very low for both male and female, but comparing just those
2 gender, the significance of p-value for male is far greater than female. This means
if a perpetrator is a male, then it would be higher chance that the shooting incident is
a murder than then when a perpetrator is a female. 

In the Age group of perpetrator, there are some typos like "1020", "224", and "940" which
are not really an age so those can be ignored. All of the age groups of perpetrator are
very low, so there is not much analysis I can make here. 

Using Null deviance and Residual deviance in the summary, I could calculate 
X^2 statistic of the model, which is a difference between those deviance.
Also, could calculate predictor value, which is the difference of
degree of freedom of Nuill and Residual deviance.
```{r calculation, echo=TRUE}
X_statistic = 17889 - 16776
degree_of_freedom = 17967-17950
X_statistic
degree_of_freedom
```
There is a Chi-Square to P-Value Calculator on-line.

https://www.statology.org/chi-square-p-value-calculator/

I put those number and got p-value of 0.000000. This is less than 0.05,
so I would conclude that this logistic regression model is highly useful.


## Bias
My bias on perpetrators might have led me to create the logistic regression model to predict
if particular perpetrators' race, age or sex relate to the murder cases. The bias is
from my background that my family used to own a convenient store in one of the cities 
with the highest crime rate. However, my bias did not affect any interpretation
on the analysis. My bias may have led me to look into certain variables but all the
analysis are based on the result from calculations and data modeling. 
